---
title: "Kalman Filter"
author:
- Eric Kalosa-Kenyon
date: "`r format(Sys.time(), '%d %B %Y')`"

abstract: I'll run you through the basics and some extensions on both the theory of the Kalman Filter and its implementation in a simple, readable R script

output:
  html_document:
    toc: yes
---

```{r message=F, echo=F}
rm(list=ls())
library("knitr")
read_chunk("./kalman.R")
```

```{r setup, echo=F}
```

# Introduction
The Kalman Filter (KF) is a method for extracting a signal from noisy
observations. Let an iterative linear process
$X_K=\{X_k | k \in [1..K] \subset \mathbf{N} \}$ be defined as follows:

$$
X_k = \Phi X_{k-1} + w_k
Z_k = H X_k + v_k
$$

# Simple example
The simplest example is to filter a constant out of noisy observations. In the
notation introduced above, $x_k\in\mathbf{R}^1$, $\Phi=1$, $w_k\aprox 0$,
$H=1$, and $V_K~WN(0,\sigma^2)$.

## Parameterization
```{r param_simple}
```

## Running the simulation
```{r simulate_simple}
```

## How does it look?
```{r plot_simple, echo=F}
```

## Performance profile
TODO

# Sources
1. Tony Lacey; "Chapter 11: Tutorial: The Kalman Filter"
2. Welch & Bishop; "An Introduction to the Kalman Filter"
