---
title: "Kalman Filter"
author:
- Eric Kalosa-Kenyon
date: "`r format(Sys.time(), '%d %B %Y')`"

abstract: I'll run you through the basics and some extensions on both the theory of the Kalman Filter and its implementation in a simple, readable R script

output:
  html_document:
    toc: yes
---

```{r message=F, echo=F}
rm(list=ls())
library("knitr")
read_chunk("./kalman.R")
```

```{r setup, echo=F}
```

# Introduction
The Kalman Filter (KF) is a method for extracting a signal from noisy
observations. Let an iterative linear process
$X_K=\{X_k | k \in [1..K] \subset \mathbf{N} \}$ be defined as follows:

(@basicdef)
$$
X_k = \Phi X_{k-1} + W_k \\
Z_k = H X_k + V_k
$$

## Filtration method
Using the model of a process defined above, we have a conception of how nature
works. That doesn't give us de facto a method to extract estimates from the
model - this is where the Kalman Filter comes in. It's a method for extracting
estimates in this context.

## Derivation details
We'll start off with a linear AR(1) process representing the change in
state as shown in the introductory section's equation (@basicdef). Assume
$W_k$ is distributed according to a centered normal distribution
$W \sim N(0, I*\sigma_w^2)$. Further, assume measurements of the state are
corrupted by measurement noise $V_k$ similarly distributed as
$V \sim N(0, I*\sigma_v^2)$. Note that the dimension of $X_k$ and $Z_k$ aren't
necessarily the same - that is, the latent space may have a different dimension
than the observation space e.g. visual observations in 2D from a dynamic system
living in 4D, 3 satellite signals each in 1D witnessing a state in 4D, etc.

Anyway, more math: we're interested in recovering the state $X_k$ as accurately
as possible.

TODO

# Simple example
The simplest example is to filter a constant out of noisy observations. In the
notation introduced above, $x_k\in\mathbf{R}^1$, $\Phi=1$, $w_k\approx 0$,
$H=1$, and $V_K \sim WN(0,\sigma^2)$. That is:

(@simpleex)
$$
x_k = x_{k-1}
Z_k = x_k + v_k
$$

We're trying to extimate $x_k$. Identically, we could define this problem as
finding $\mu$ from the process $Z_K$:

$$
Z_K \sim WN(\mu, I_K\sigma^2)
$$

## Parameterization
```{r param_simple}
```

## Running the simulation
```{r simulate_simple}
```

## How does it look?
### The filter itself
```{r plot_simple, echo=F}
```

### The error profile
```{r plot_error_profile_1d, echo=F}
```
  
It appears that the error becomes bounded extremely quickly, within 10 or so
observations.

# Adding some movement

# 2D Filter
Coming soon

# Sources
1. Tony Lacey; "[Chapter 11: Tutorial: The Kalman Filter](http://web.mit.edu/kirtley/kirtley/binlustuff/literature/control/Kalman%20filter.pdf)"; MIT
2. Welch & Bishop (2006); "[An Introduction to the Kalman Filter](https://courses.cs.washington.edu/courses/cse571/03wi/notes/welch-bishop-tutorial.pdf)"; UNC Chapel Hill
